---
title: "P8131_HW4"
output: html_document
date: "2023-03-09"
---

```{r setup, include=FALSE}
library(tidyverse)
library(survey)
library(nnet)
library(MASS)
```

```{r}
housing = 
  c(65, 34, 130, 141, 67, 130, 54, 47, 76, 116, 48, 105, 100, 100, 111, 191, 62, 104) %>%
  array(c(2, 3, 3),
        dimnames = list(contact = c("Low", "High"),
                        type = c("Tower", "Apartment", "House"),
                        satisfaction = c("sat.low", "sat.medium", "sat.high")))
ftable(housing)
```
### QUESTION 1

#### Pair-wise associations 1) contact with other residents:

```{r}
mar_con = margin.table(housing, margin=c(1, 3))
prop.table(mar_con, margin = 1)
chisq.test(mar_con)
```
There seems to be no association between contact with others and levels of satisfaction. Chi-square test shows that under the null hypothesis of no association, the p-value is 0.076 > 0.05 , we fail to reject the null and claim that there’s no association between level of contact and level of satisfaction.

#### Pair-wise associations 2) type of housing:

```{r}
mar_type = margin.table(housing, margin=c(2, 3))
prop.table(mar_type, margin = 1)
chisq.test(mar_type)
```
It seems that those live in tower blocks have the highest percent of high satisfaction and lowest percent of low satisfaction, while those live in houses have the lowest percent of high satisfaction and highest percent of low satisfaction. Chi-square test shows that under the null hypothesis of no association, the p-value is 7.369e-07 << 0.05 , we reject the null and claim that there’s association between type of housing and level of satisfaction.

### QUESTION 2

#### Fit a multinomial model:

```{r results='hide'}
housing_df = 
  as.data.frame.table(housing) %>% 
  pivot_wider(
    names_from = "satisfaction",
    values_from = "Freq") %>% 
  mutate(type = factor(type, levels = c('Apartment', 'House', 'Tower')))

m.mult <- multinom(cbind(sat.low, sat.medium, sat.high) ~ contact + type, data=housing_df)
```

```{r}
summary(m.mult)
```
#### Pattern in the associations and odds ratios with 95% confidence intervals:

The multinomial model for this data can be written as: 

log(πj/πi) = β0j + β1j·X1 + β2j·X2 + β3j·X3 With X’s stand for:

* X1=0, when contact is low; X1=1, when contact is high;

* X2=1, when live in houses; X3=1, when live in tower blocks; X2=X3=0, when live in apartments.

```{r}
exp(coef(m.mult))
exp(confint(m.mult))
```
* β12: Under the same housing types, the OR of satisfaction level between medium and low increased by 1.34 (95% CI: 1.0418307-1.7349522) when the contact changes from low to high.

* β22: Under the same contact levels, the OR of satisfaction level between medium and low increased by 1.07 (95% CI: 0.8088711-1.4211574) when the housing type changes from apartment to house, as the confidence includes 1, this association is not significant.

* β32: Under the same contact levels, the OR of satisfaction level between medium and low increased by 1.51 (95% CI: 1.0735958-2.1011828) when the housing type changes from apartment to tower block.

* β13: Under the same housing types, the OR of satisfaction level between high and low increased by 1.39 (95% CI: 1.1014013-1.7504412) when the contact changes from low to high.

* β23: Under the same contact levels, the OR of satisfaction level between high and low decreased by 0.74 (95% CI: 0.5661217-0.9616618) when the housing type changes from apartment to house.

* β33: Under the same contact levels, the OR of satisfaction level between high and low increased by 1.90 (95% CI: 1.4154471-2.5490928) when the housing type changes from apartment to tower block.

#### Goodness-of-fit:

```{r}
# pearson residuals
pihat=predict(m.mult,housing_df,type='probs')
m=rowSums(housing_df[,3:5])
res.pearson=(housing_df[,3:5]-pihat*m)/sqrt(pihat*m)
# Generalized Pearson Chisq Stat
G.stat=sum(res.pearson^2)
G.stat
pval=1-pchisq(G.stat,df=(6-4)*(3-1))
pval
# deviance
D.stat=sum(2*housing_df[,3:5]*log(housing_df[,3:5]/(m*pihat)))
D.stat
```
The p-value for eneralized pearson's chi-squared test is 0.14 and is greater than 0.05, indicating fail to reject the null and the fit is good. The pearson chisq-stat is 6.93 and deviance 6.89, both with 4 degrees of freedom.

#### Interaction:

```{r results='hide'}
m.mult_in <- multinom(cbind(sat.low, sat.medium, sat.high) ~ (contact + type)^2, data=housing_df)
```


```{r}
summary(m.mult_in)
dev = m.mult_in$deviance - m.mult$deviance
1-pchisq(dev, df = 2)
```
We used deviance for model selection, the p-value for chi-square test is greater than 0.05, we failed to reject the null and claim that the smaller model is better and there is no interaction.

### QUESTION 3

#### Fit a proportional odds model:

```{r results='hide'}
housing_df2 = 
  as.data.frame.table(housing)%>% 
  mutate(type = factor(type, levels = c('Apartment', 'House', 'Tower')))
m.mo = polr(satisfaction~contact+type,data=housing_df2,weights=Freq)
```

```{r}
summary(m.mo)
```

```{r}
exp(coef(m.mo))
exp(confint(m.mo))
```
#### Model intepretation: 

The proportional odds model for this data can be written as: 

log( γj/1- γi) = β0j + β1j·X1 + β2j·X2 + β3j·X3 with X’s stand for:

* X1=0, when contact is low; X1=1, when contact is high;

* X2=1, when live in houses; X3=1, when live in tower blocks; X2=X3=0, when live in apartments.


